---
title: "Lab4_Demo"
author: "Mateo Robbins"
date: "2024-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)     # for data wrangling
library(ggplot2)   # for awesome plotting
library(rsample)   # for data splitting
library(recipes)   # data preprocessing
library(caret)     # for logistic regression modeling
```
Let's explore how employee income and overtime hours worked affect likelihood of employee attrition.  Any predictions?

```{r}
#
data("attrition", package = "modeldata")

df <- attrition %>% 
  mutate_if(is.ordered, factor, ordered = FALSE)

# Create training (70%) and test (30%) sets for the 
# rsample::attrition data.
set.seed(123)  # for reproducibility 
churn_split <- initial_split(df, prop = 0.7)
churn_train <- training(churn_split)
churn_test  <- testing(churn_split)
```

Let's set up set up our recipes now for preprocessing. 
```{r recipe}
#specify and prep recipe
churn_rec <- recipe(Attrition ~ ., data = churn_train) %>% # set recipe
  step_integer(Attrition, zero_based = T) %>% # make attrition into 0/1 form
  prep(churn_train)

#bake recipe with training data
churn_baked_train <- bake(churn_rec, # object with recipe
                          new_data = churn_train) # df that the preprocessing will be applied
```

# Created 2 models to predict attrition

```{r specify_models_glm}
#MonthlyIncome
model_inc <- glm(data = churn_baked_train, 
                 Attrition ~ MonthlyIncome, 
                 family = "binomial") # because outcome variable is binary
  
#OverTime
model_time <- glm(data = churn_baked_train, 
                 Attrition ~ OverTime, 
                 family = "binomial") # because outcome variable is binary
```


```{r tidy_model_objs}
broom::tidy(model_inc)
broom::tidy(model_time)

```
Hard to interpret the coefficients. How does this impact the likelihood of attrition? We can exponentiate the coefficients, so that we can understand the difference in the predictor variables. 

The odds of an employee attriting, according to the output of `exp(coef(model_inc))` increase by 0.999 for every additional dollar. Since this is less than 1, the odds are decreasing. This is because you are multiplying the `exp`. 

For the overtime, the odds of an employee attriting increases by 3.48 when an employee has worked overtime. This means that when employees are working overtime, the odds of them leaving increase.   

```{r exp_coefs}
#exponentiate the coefficients from model objects for interpretation. Gives us changes in odds of attrition
exp(coef(model_inc))

exp(coef(model_time))
```

# See how model performs on test data 

```{r recode_attrition_test}
# apply recipe to test data
churn_baked_test <- bake(churn_rec, 
                         new_data = churn_test)
```

```{r plot_income_attrition}
ggplot(churn_baked_test, aes(x = MonthlyIncome, y = Attrition)) +
  geom_point() +
  stat_smooth(method = "glm", 
              se = T,
              method.args = list(family = "binomial"))

```

We can add more predictors, creating a multiple logistic regression model

```{r mult_log_regression}
model_both <- glm(Attrition ~ MonthlyIncome + OverTime, family = "binomial",
                  data = churn_train)

broom::tidy(model_both)
```

```{r}
# representing what model_both is representing 
ggplot(churn_baked_test, 
       aes(x = MonthlyIncome, y = Attrition, color = OverTime)) +
  geom_point(alpha = 0.5) +
  stat_smooth(method = "glm", 
              se = F,
              method.args = list(family = "binomial"))
```


