---
title: "Lab5_Demo"
author: "Mateo Robbins"
date: "2024-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)    
library(ggplot2) #great plots
library(rsample)  #data splitting 
library(recipes) #data preprocessing
library(skimr) #data exploration
library(tidymodels) #re-entering tidymodel mode
library(kknn) #knn modeling
```

###k-nearest neighbor in tidymodels

## Data

```{r data}
data(attrition)
churn <- attrition %>% mutate_if(is.ordered, .funs = factor, ordered = F) 
#skim(churn) run in console
```

Not doing the data exploration here in the interest of time and since we are familiar with this dataset.

```{r initial_split}
set.seed(808)
#initial split of data, default 75/25
churn_split <- initial_split(churn)
churn_test <- testing(churn_split)
churn_train <- training(churn_split)
```

We need to create a recipe and do the preprocessing by dummy coding the nominal variables and standardizing the numeric variables.

```{r recipe}
#preprocessing
knn_rec <- recipe(Attrition ~ ., data = churn_train) %>% 
  
  # apply dummy coding to all nominal variables but NOT the outcome variables
  step_dummy(all_nominal(), -all_outcomes(),
             
             # factors/level becomes column, and encoded as either 1 or 0
             one_hot = T) %>%  # type of dummy encoding; what is being applied to the variables
  
  # need to normalize the distance / scale of the variables to get one the same scale
  step_normalize(all_numeric(), -all_outcomes()) %>% 
  prep()
  
# apply the recipe to the training data 
baked_train <- bake(knn_rec, churn_train)
```

Recall: if you want to see the what the recipe is doing to your data, you can first prep() the recipe to estimate the parameters needed for each step and then bake(new_data = NULL) to pull out the training data with those steps applied.

Now the recipe is ready to be applied to the test data.

```{r bake_test}
baked_test <- bake(knn_rec, churn_test)
```

##Specify the k-nearest neighbor model

```{r knn_spec}
# how many neighbors to look at 
knn_spec <- parsnip::nearest_neighbor(neighbors = 5) %>% 
  set_engine("kknn") %>% 
  
  # is it a regression or classification task
  set_mode("classification") 
```

## We are only using `fit()` because we are using 1 single value and testing it. This step isn't needed once we do the workflow(). 
```{r knn_fit}
knn_fit <- knn_spec %>% # give specificatiion
  fit(Attrition ~ ., data = churn_train) # outcome variable


# read response
knn_fit
```

The minimal miscalssification gives the error rate. To get the accuracy to 1 - minimal_classification.

```{r cv}
set.seed(808)
# 5-fold CV on the training dataset (instead of 10 for in-class demo)
cv_folds <- churn_train %>% 
  vfold_cv(v = 5)

```

We now have a recipe for processing the data, a model specification, and CV splits for the training data.

Let's put it all together in a workflow.

```{r knn_workflow}
knn_workflow <- workflow() %>% 
  add_model(knn_spec) %>% 
  add_recipe(knn_rec)

```

Let's fit the resamples and carry out the cross-validation
```{r knn_res}
knn_res <- 
  knn_workflow %>% 
  fit_resamples(
    resamples = cv_folds, # defined above 
    control = control_resamples(save_pred = T)
  )
```

```{r}
# Check the performance
knn_res %>% collect_metrics()
```
Ran the model 5 times over each fold, and calculated the accuracy and averaged them over the 5 runs. The values for the performance metrics is here. This is the accuracy using the value of neighbors that we specified above (5). 

Let's find the best value of k
```{r spec_with_tuning}
# Define our KNN model with tuning

# set the neigbors to tuning so that it selects the best value of neighbors to increase accuracy
knn_spec_tune <- nearest_neighbor(neighbors = tune()) %>% 
  set_mode("classification") %>% 
  set_engine("kknn")

```

```{r wf_knn_tune}
# Define a new workflow
wf_knn_tune <- workflow() %>% 
  add_model(knn_spec_tune) %>% 
  add_recipe(knn_rec)
```

This time before we fit the model we need to tell R which values to try for the parameter that we're tuning.

To tune our hyperparameter(s), we will use the tune_grid() function (instead of the fit() or fit_resamples() functions).

This tune_grid() is similar to fit_resamples() except that it takes an additional argument: grid. We will pass the possible values of our hyperparameter(s) to this grid argument, and it will evaluate each fold of our sample on each set of hyperparameters passed to grid.

We'll explore a few values of k: (1,5,10, 20, 30, 40, 50, 60, 70, 80, 90, 100)

**Runs the model 60 times, because we have 12 values of n, and its running cv 5 times on each. 5 x 12 is 60.**
```{r fit_knn_cv}
# Fit the workflow on our predefined folds and a grid of hyperparameters
fit_knn_cv <- wf_knn_tune %>% 
  tune_grid(
    cv_folds,
    grid = data.frame(neighbors = c(1, 5, seq(10, 100, 10)))
  )

# Check the performance with collect_metrics()
fit_knn_cv %>% collect_metrics()
```

And finally, we will predict.

Use finalize_workflow() function wants (1) your initial workflow and (2) your best model.

```{r final_wf}
# The final workflow for our KNN model. Finalize_workflow takes a workflow and a set of parameters.  In this case, that set is just the best value of k
final_wf <- wf_knn_tune %>% 
  finalize_workflow(select_best(fit_knn_cv, metric = "accuracy"))

# Check out the final workflow object.  Choosing accuracy for interpretability in this simple binary context
final_wf
```

```{r final_fit}
# Fitting our final workflow
final_fit <- final_wf %>% 
  fit(data = churn_train)
# Examine the final workflow
final_fit
```

And finally, we can predict onto the testing dataset.

```{r churn_pred}
churn_pred <- final_fit %>% 
  predict(new_data = churn_test)

churn_pred
```

There's a better way! You can pass your final workflow (workflow plus the best model) to the last_fit() function along with your initial split (for us: churn_split) to both (a) fit your final model on your full training dataset and (b) make predictions onto the testing dataset (defined in your initial split object).

This last_fit() approach streamlines your work (combining steps) and also lets you easily collect metrics using the collect_metrics() function

**By using last_fit() we are skipping the fit() to the test data**

```{r last_fit}
# Write over 'final_fit' with this last_fit() approach
final_fit <- final_wf %>% 
  
  # fit on the training data, predict on the test data 
  last_fit(churn_split)

# Collect metrics on the test data!
final_fit %>%  collect_metrics()
```
