---
title: "Lab4"
author: "Luna Herschenfeld-Catalan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(skimr)
library(tidymodels)
library(caret)
# Load the corrplot package
library(corrplot)
library(sjPlot)
```

## Lab 4: Fire and Tree Mortality

The database we'll be working with today includes 36066 observations of individual trees involved in prescribed fires and wildfires occurring over 35 years, from 1981 to 2016. It is a subset of a larger fire and tree mortality database from the US Forest Service (see data description for the full database here: [link](https://www.nature.com/articles/s41597-020-0522-7#Sec10)). Our goal today is to predict the likelihood of tree mortality after a fire.

### Data Exploration

Outcome variable: *yr1status* = tree status (0=alive, 1=dead) assessed one year post-fire.

Predictors: *YrFireName, Species, Genus_species, DBH_cm, CVS_percent, BCHM_m, BTL* (Information on these variables available in the database metadata ([link](https://www.fs.usda.gov/rds/archive/products/RDS-2020-0001-2/_metadata_RDS-2020-0001-2.html))).

```{r}
trees_dat<- read_csv(file = "https://raw.githubusercontent.com/MaRo406/eds-232-machine-learning/main/data/trees-dat.csv")
```
### Data Splitting
> Question 1: Recode all the predictors to a zero_based integer form
> Question 2: Create trees_training (70%) and trees_test (30%) splits for the modeling

```{r}
set.seed(23)

trees_split <- initial_split(trees_dat, prop = 0.7) # create training and test data
trees_train <- training(trees_split)
trees_test <- testing(trees_split)

trees_recipe <- recipe(yr1status ~ ., # have status as outcome variable
                      data = trees_train) %>%  # set recipe
  step_integer(all_string(), # make all character predictors into integers
               zero_based = T) %>% 
  prep(trees_train)

trees_baked_train <- bake(trees_recipe, 
                          new_data = trees_train)
```

> Question 3: How many observations are we using for training with this split?

The training data contains 25,246 observations. 

```{r}
dim(trees_train)
```


### Simple Logistic Regression 

Let's start our modeling effort with some simple models: one predictor and one outcome each.

> Question 4: Choose the three predictors that most highly correlate with our outcome variable for further investigation.

```{r}
# Obtain correlation matrix
corr_mat <- cor(trees_baked_train)

# Make a correlation plot between the variables
corrplot(corr_mat, method = "shade", shade.col = NA,
         tl.col = "black",
         tl.srt = 45,
         addCoef.col = "black",
         cl.pos = "n",
         order = "original")

```

Based on our correlation matrix, the variables that are most correlated with the `yr1status` are `CVS_percent`, `BCHM_m`, and `DBH_cm`.

> Question 5: Use glm() to fit three simple logistic regression models, one for each of the predictors you identified.

```{r}
mod1 <- glm(yr1status ~ CVS_percent, data = trees_train, family = "binomial")

mod2 <- glm(yr1status ~ BCHM_m, data = trees_train, family = "binomial")

mod3 <- glm(yr1status ~ DBH_cm, data = trees_train, family = "binomial")
```


### Interpret the Coefficients 

We aren't always interested in or able to interpret the model coefficients in a machine learning task. Often predictive accuracy is all we care about.

> Question 6: That said, take a stab at interpreting our model coefficients now.

```{r}
tab_model(mod1)
tab_model(mod2)
tab_model(mod3)
```

For the first model, the coefficient for CVS percent is 1.08, with a p-value of less than 0.001. This can be understood that we are over 99% confidence that for every 1 percent increase in crown volume scorched (CVS), the odds of the tree status 1 year post fire being dead increases by 1.08. For the second model, the coefficient for BCHM_m is 1.24, with a p-value of less than 0.001. This can be understood that we are over 99% confident that for every 1 meter vertical increase of maximum bark char height from the ground, the odds of the tree status 1 year post fire being dead increases by 1.24. For the third model, the coefficient for DBH_cm is 0.94, with a p-value of less than 0.001. This can be understood that we are over 99% confident that for every 1 centimeter increase of diameter, the odds of the tree status 1 year post fire being dead increases by 0.94. Since this is less than 1, the odds are decreasing as the diameter of the tree increases. 

> Question 7: Now let's visualize the results from these models. Plot the fit to the training data of each model.

```{r}
ggplot(trees_baked_train, aes(x = CVS_percent, y = yr1status)) +
  geom_point() +
  stat_smooth(method = "glm", 
              se = T,
              method.args = list(family = "binomial"))

ggplot(trees_baked_train, aes(x = BCHM_m, y = yr1status)) +
  geom_point() +
  stat_smooth(method = "glm", 
              se = T,
              method.args = list(family = "binomial"))

ggplot(trees_baked_train, aes(x = DBH_cm, y = yr1status)) +
  geom_point() +
  stat_smooth(method = "glm", 
              se = T,
              method.args = list(family = "binomial"))
```


### Multiple Logistic Regression

Let's not limit ourselves to a single-predictor model. More predictors might lead to better model performance.

> Question 8: Use glm() to fit a multiple logistic regression called "logistic_full", with all three of the predictors included. Which of these are significant in the resulting model?

All three of these variables are significant in the resulting model. 

```{r}
logistic_full <- glm(yr1status ~ CVS_percent + BCHM_m + DBH_cm, 
                     data = trees_train, 
                     family = "binomial") 

summary(logistic_full)

broom::tidy(logistic_full)

```


### Estimate Model Accuracy

Now we want to estimate our model's generalizability using resampling.

> Question 9: Use cross validation to assess model accuracy. Use caret::train() to fit four 10-fold cross-validated models (cv_model1, cv_model2, cv_model3, cv_model4) that correspond to each of the four models we've fit so far: three simple logistic regression models corresponding to each of the three key predictors (CVS_percent, DBH_cm, BCHM_m) and a multiple logistic regression model that combines all three predictors.

```{r eval = FALSE}
trees_train$yr1status <- as.factor(trees_train$yr1status)

# create subsets for cross validation 
ctrl <- trainControl(method = "cv",
                     number = 10) # default number of k-folds

set.seed(123)
cv_model1 <- caret::train(yr1status ~ CVS_percent, 
                          data = trees_train,
                          method = "glm",
                          family = "binomial",
                          trControl = ctrl)
set.seed(123)
cv_model2 <- caret::train(yr1status ~ BCHM_m, 
                          data = trees_train,
                          method = "glm",
                          family = "binomial",
                          trControl = ctrl)
set.seed(123)
cv_model3 <- caret::train(yr1status ~ DBH_cm, 
                          data = trees_train,
                          method = "glm",
                          family = "binomial",
                          trControl = ctrl)
set.seed(123)
cv_model4 <- caret::train(yr1status ~ CVS_percent + BCHM_m + DBH_cm, 
                          data = trees_train,
                          method = "glm",
                          family = "binomial",
                          trControl = ctrl)
```


> Question 10: Use caret::resamples() to extract then compare the classification accuracy for each model. (Hint: resamples() wont give you what you need unless you convert the outcome variable to factor form). Which model has the highest accuracy?

```{r}

# extract sample performance 
summary(caret::resamples(
  
  list(
    # model on the left is the one WE made
    cv_model1 = cv_model1, 
    cv_model2 = cv_model2, 
    cv_model3 = cv_model3,
    cv_model4 = cv_model4
  ))
  )$statistic$Accuracy

```


Let's move forward with this single most accurate model.

> Question 11: Compute the confusion matrix and overall fraction of correct predictions by the model.

```{r}
predicted <- predict(cv_model4, trees_train)

# create confusion matrix
confusionMatrix(data = predicted, 
                reference = trees_train$yr1status)
```

> Question 12: Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.
>
> Question 13: What is the overall accuracy of the model? How is this calculated?

### Test Final Model

Alright, now we'll take our most accurate model and make predictions on some unseen data (the test data).

> Question 14: Now that we have identified our best model, evaluate it by running a prediction on the test data, trees_test.

```{r}
# fore generating test metrics
# predicted_test <- predict(cv_model4, trees_test) %>% 
#   bind_cols(trees_test) %>% 
#   rename("pred" = "...1") %>% 
#   mutate(pred = as.numeric(pred))
# 
# test_metrics <- predicted_test %>% 
#   metrics(yr1status, pred)
# 
# test_metrics

# for generating confusion matrix
predicted_test_df <- predict(cv_model4, trees_test)

# create confusion matrix
confusionMatrix(data = predicted_test_df, 
                reference = as.factor(trees_test$yr1status))

```


> Question 15: How does the accuracy of this final model on the test data compare to its cross validation accuracy? Do you find this to be surprising? Why or why not?
