---
title: "Lab 8"
author: "Luna Herschenfeld-Catalán"
date: "2024-03-06"
output: html_document
---

## Forest Cover Classification with SVM

In this week's lab we are exploring the use of Support Vector Machines for multi-class classification. Specifically, you will be using cartographic variables to predict forest cover type (7 types).

Natural resource managers responsible for developing ecosystem management strategies require basic descriptive information including inventory data like forest cover type for forested lands to support their decision-making processes. However, managers generally do not have this type of data for in-holdings or neighboring lands that are outside their immediate jurisdiction. One method of obtaining this information is through the use of predictive models.

You task is build both an SVM and a random forest model and compare their performance on accuracy and computation time.

1.  The data is available here: <https://ucsb.box.com/s/ai5ost029enlguqyyn04bnlfaqmp8kn4>
```{r setup}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidymodels)
library(tidyverse)
library(dplyr)
library(kernlab)
library(tidyclust)

set.seed(47)
```


Explore the data.

```{r}
covtype <- read_csv(here::here("labs", "data", "covtype_sample.csv")) %>% 
  janitor::clean_names() %>% 
  mutate(cover_type = as.factor(cover_type)) %>% 
  mutate(cover_type = as.factor(cover_type)) %>%  
  mutate(across(6:54, as.factor)) %>%   # convert binary to initial
  select(-soil_type_15, -soil_type_37, -soil_type_38, -soil_type_39, -soil_type_40)
  #select_if(function(col) length(unique(col))>1) %>%  # remove columns with only 0 in them

```

-   What kinds of features are we working with?

**Some of the features are continuous and some are categorical.**

-   Does anything stand out that will affect you modeling choices?

Hint: Pay special attention to the distribution of the outcome variable across the classes.

**Since there are both categorical and continuous variables, in our recipe we need to think critically about whether we are using a linear engine or a classification engine in the workflow. The `soil_type` variables are all categorical though, and it is coding for presence. When we plot the distribution of the `cover_type` which is the outcome variable, we also see that there is a large bias in the dataset. Most of the points are the `cover_type` 1 and 2, and `cover_type` 4 is very underrepresented.**

```{r}
ggplot(covtype) +
  geom_histogram(aes(x = cover_type), stat="count") +
  labs(title = "Distribution of Cover Types in Dataset") +
  theme_minimal() 
```


2.  Create the recipe and carry out any necessary preprocessing. Can you use the same recipe for both models?

**We can use the same recipe for both models because we are treating the data based on the attributes**

```{r}
# split the data 
cov_split <- initial_split(covtype)
cov_train <- training(cov_split)
cov_test <- testing(cov_split)

# split the data to only keep 10% because svm thinks its redundant
cov_split_svm <- initial_split(cov_train, 
                               strata = cover_type,
                               prop = 0.1)
cov_train_svm <- training(cov_split_svm)
cov_test_svm <- testing(cov_split_svm)

# create recipe with training data 
cov_recipe <- recipe(cover_type ~ ., data = cov_train) %>% 
  step_dummy(all_nominal_predictors(), # categorical predictors
             one_hot = TRUE) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_center(all_numeric_predictors()) %>% 
  step_scale(all_numeric_predictors()) # all the numeric variables

#centering and scale
#scale that up (look at the best parameters for that so that tuning isnt trying too much - use small data set)

cov_recipe_svm <- recipe(cover_type ~ ., data = cov_train_svm) %>% 
  step_dummy(all_nominal_predictors(), # categorical predictors
             one_hot = TRUE) %>% 
  step_normalize(all_numeric_predictors())
```


3.  Create the folds for cross-validation.
```{r}
# create the folds 
vfolds <- vfold_cv(cov_train, strata = cover_type)
```


4.  Tune the models. Choose appropriate parameters and grids. If the computational costs of tuning given your strategy are prohibitive, how might you work around this?

# SVM
```{r}
# create clustering 
svm_rbf_spec <- svm_rbf() %>% 
  set_mode("classification") %>% 
  set_engine("kernlab")

svm_rbf_fit <- svm_rbf_spec %>% 
  fit(cover_type ~ ., data = cov_train_svm)

svm_rbf_augment <- augment(svm_rbf_fit, new_data = cov_test_svm)

# confusion matrix
svm_rbf_augment %>% 
  conf_mat(truth = cover_type, estimate = .pred_class)

# roc curve
svm_rbf_augment %>%
  roc_curve(truth = cover_type, .pred_1, .pred_2, .pred_3, .pred_4, .pred_5, .pred_6, .pred_7) %>% 
  autoplot()

# roc curve
svm_rbf_augment %>%
  roc_auc(truth = cover_type, .pred_1, .pred_2, .pred_3, .pred_4, .pred_5, .pred_6, .pred_7)

```


```{r eval = FALSE}

# create clustering 
svm_spec <- svm_poly(degree = 1, cost = tune()) %>% 
  set_mode("classification") %>% 
  set_engine("kernlab")

# create workflow
svm_wf <- workflow() %>% 
  add_model(svm_spec %>% 
              set_args(cost = tune())) %>% 
  add_recipe(cov_recipe_svm)
  

# create grid
param_grid <- grid_regular(cost(), 
                           levels = 10) # number of different values for cost() that we want to try

# tune the cluster
# recipe with the step_normalize() did not work so I removed it
tune_res <- tune_grid(
  svm_wf, # add workflow
  resamples = vfolds, # add folds
  grid = param_grid # add grid
)

autoplot(tune_res)
```


# Random Forest
```{r eval = FALSE}
rf_spec <- rand_forest(mtry = tune(), 
                        trees = tune()) %>%
  set_engine("ranger") %>% 
  set_mode("classification") # the output is cover type

# create workflow with feature recipe
rf_workflow <- workflow() %>% 
  add_model(rf_spec) %>% 
  add_recipe(cov_recipe_svm)

# this is taking a r
rf_tune <- rf_workflow %>% 
  tune_grid(
    resamples = vfolds, # add folds
    grid = 2 # number of combos of mtry and trees
  )

```


5.  Conduct final predictions for both models and compare their prediction performances and computation costs from part 4.

-   Which type of model do you think is better for this task?
-   Why do you speculate this is the case?
